---
title: "CA Notify Adoption Univariable & Multivariable Analysis"
author: "Kevin Nguyen, Ravi Goyal, Mingjia Zhu, Gary Adamos, Rachel Kunowski"
date: "5/13/2022"
output: html_document
---
### Loading Libraries
```{r, message=FALSE, warning=FALSE}
library(googlesheets4)
library(glmnet)
library(ggiraph)
library(ggiraphExtra)
library(randomForest)
library(ggplot2)
library(dplyr)
set.seed(1)
```

### Loading Data
```{r, include = FALSE}
gs4_deauth()
ENXadoption = read_sheet(ss = "hidden")
ENXadoption_matrix = as.matrix(ENXadoption[c(3:22,28)])
ENXadoption_demographics_matrix = as.matrix(ENXadoption[c(3:16)])
```

## Univariable Analysis
```{r}
bachelors_morelm = lm(adoption~bachelors_more, data = ENXadoption)
summary(bachelors_morelm)
ggPredict(bachelors_morelm, se = TRUE)

whitelm = lm(adoption~white, data = ENXadoption)
summary(whitelm)
ggPredict(whitelm, se = TRUE)

blacklm = lm(adoption~black, data = ENXadoption)
summary(blacklm)
ggPredict(blacklm, se = TRUE)

asianlm = lm(adoption~asian, data = ENXadoption)
summary(asianlm)
ggPredict(asianlm, se = TRUE)

other_racelm = lm(adoption~other_race, data = ENXadoption)
summary(other_racelm)
ggPredict(other_racelm, se = TRUE)

hispaniclm = lm(adoption~hispanic, data = ENXadoption)
summary(hispaniclm)
ggPredict(hispaniclm, se = TRUE)

income100k_morelm = lm(adoption~income100k_more, data = ENXadoption)
summary(income100k_morelm)
ggPredict(income100k_morelm, se = TRUE)

householdsize3_morelm = lm(adoption~householdsize3_more, data = ENXadoption)
summary(householdsize3_morelm)
ggPredict(householdsize3_morelm, se = TRUE)

democratlm = lm(adoption~democrat, data = ENXadoption)
summary(democratlm)
ggPredict(democratlm, se = TRUE)

republicanlm = lm(adoption~republican, data = ENXadoption)
summary(republicanlm)
ggPredict(republicanlm, se = TRUE)

unaffiliated_partylm = lm(adoption~unaffiliated_party, data = ENXadoption)
summary(unaffiliated_partylm)
ggPredict(unaffiliated_partylm, se = TRUE)

other_partylm = lm(adoption~other_party, data = ENXadoption)
summary(other_partylm)
ggPredict(other_partylm, se = TRUE)

olderthan60lm = lm(adoption~olderthan60, data = ENXadoption)
summary(olderthan60lm)
ggPredict(olderthan60lm, se = TRUE)

death_ratelm = lm(adoption~death_rate, data = ENXadoption)
summary(death_ratelm)
ggPredict(death_ratelm, se = TRUE)

caseslm = lm(adoption~cases, data = ENXadoption)
summary(caseslm)
ggPredict(caseslm, se = TRUE)

deathslm = lm(adoption~deaths, data = ENXadoption)
summary(deathslm)
ggPredict(deathslm, se = TRUE)

hospitalizationslm = lm(adoption~hospitalizations, data = ENXadoption)
summary(hospitalizationslm)
ggPredict(hospitalizationslm, se = TRUE)

maskfrequently_morelm = lm(adoption~maskfrequently_more, data = ENXadoption)
summary(maskfrequently_morelm)
ggPredict(maskfrequently_morelm, se = TRUE)

at_least_partially_vaccinatedlm = lm(adoption~at_least_partially_vaccinated, data = ENXadoption)
summary(at_least_partially_vaccinatedlm)
ggPredict(at_least_partially_vaccinatedlm, se = TRUE)

fully_vaccinatedlm = lm(adoption~fully_vaccinated, data = ENXadoption)
summary(fully_vaccinatedlm)
ggPredict(fully_vaccinatedlm, se = TRUE)

residentiallm = lm(adoption~residential, data = ENXadoption)
summary(residentiallm)
ggPredict(residentiallm, se = TRUE)
```

# Demographics  
## Multivariable Analysis 
### Univariable Analysis p<0.1 [also less than 0.05]
```{r}
demographics_p_less_0.1 = lm(adoption~bachelors_more + white + black + asian + income100k_more + democrat + republican + unaffiliated_party + other_party + death_rate, 
                 data = ENXadoption)
summary(demographics_p_less_0.1)
```

### Selection
```{r, include=FALSE}
all_selection_demographics = lm(adoption~bachelors_more + white + black + asian + other_race + hispanic + income100k_more + householdsize3_more + democrat + republican + unaffiliated_party + other_party + olderthan60 + death_rate, 
                 data = ENXadoption)

intercept_selection_demographics = lm(adoption~1, data = ENXadoption)
```
#### Backward Selection
```{r}
backward_selection_demographics = step(all_selection_demographics, direction = "backward", scope = formula(intercept_selection_demographics), trace = 0)
summary(backward_selection_demographics)
```
#### Forward Selection
```{r}
forward_selection_demographics = step(intercept_selection_demographics, direction = "forward", scope = formula(all_selection_demographics), trace = 0)
summary(forward_selection_demographics)
```
#### Stepwise Selection
```{r}
stepwise_selection_demographics = step(intercept_selection_demographics, direction = "both", scope = formula(all_selection_demographics), trace = 0)
summary(stepwise_selection_demographics)
```

## LASSO Analysis
```{r}
LASSO_demographics = glmnet(x = ENXadoption_demographics_matrix, y = ENXadoption$adoption)
plot(LASSO_demographics, xvar = "lambda", label = TRUE)
```

## Random Forest Analysis
```{r}
RForest_demographics = randomForest(x = ENXadoption[3:16], y = ENXadoption$adoption, importance = TRUE)
print(RForest_demographics)
round(importance(RForest_demographics), 2)
varImpPlot(RForest_demographics)
```

# Total  
## Multivariable Analysis 
### Univariable Analysis p<0.1 [also less than 0.05]
```{r}
p_less_0.1 = lm(adoption~bachelors_more + white + black + asian + income100k_more + democrat + republican + unaffiliated_party + other_party + death_rate + maskfrequently_more + at_least_partially_vaccinated + fully_vaccinated + residential, 
                 data = ENXadoption)
summary(p_less_0.1)
```

### Selection
```{r, include=FALSE}
all_selection = lm(adoption~bachelors_more + white + black + asian + other_race + hispanic + income100k_more + householdsize3_more + democrat + republican + unaffiliated_party + other_party + olderthan60 + death_rate + cases + deaths + hospitalizations + maskfrequently_more + at_least_partially_vaccinated + fully_vaccinated + residential, 
                 data = ENXadoption)

intercept_selection = lm(adoption~1, data = ENXadoption)
```
#### Backward Selection
```{r}
backward_selection = step(all_selection, direction = "backward", scope = formula(intercept_selection), trace = 0)
summary(backward_selection)
```
#### Forward Selection
```{r}
forward_selection = step(intercept_selection, direction = "forward", scope = formula(all_selection), trace = 0)
summary(forward_selection)
```
#### Stepwise Selection
```{r}
stepwise_selection = step(intercept_selection, direction = "both", scope = formula(all_selection), trace = 0)
summary(stepwise_selection)
```

## LASSO Analysis

```{r}
LASSO_total = glmnet(x = ENXadoption_matrix, y = ENXadoption$adoption)
plot(LASSO_total, xvar = "lambda", label = TRUE)
```

## Random Forest Analysis
```{r}
RForest_total = randomForest(x = ENXadoption[c(3:20,22,28)], y = ENXadoption$adoption, importance = TRUE)
print(RForest_total)
round(importance(RForest_total), 2)
varImpPlot(RForest_total)

RFtotal = as.data.frame(varImpPlot(RForest_total))
RFtotal$varnames = rownames(RFtotal)
rownames(RFtotal) = NULL
RFtotal$variables = c("Bachelor’s Degree or Higher", "White", "Black", "Asian", "Other Race", "Hispanic", "$100,000 or More", "3 People or More", "Democratic", "Republican", "No Affiliation", "Other Party", "Over 60 Years Old", "Death Rates (Pre-COVID-19)", "Cases", "Deaths", "Hospitalizations", "Masking", "Fully Vaccinated", "Mobility")

ggplot(RFtotal, aes(x=reorder(variables, IncNodePurity), weight=IncNodePurity, fill="red")) +
  geom_bar() + ylab("Increase in Node Purity") + xlab("Variable Name") + theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) +
  geom_hline(yintercept = 0.019, color ="blue") + coord_flip()

ggplot(RFtotal, aes(x=reorder(variables, `%IncMSE`), weight=`%IncMSE`, fill="red")) + 
  geom_bar() + ylab("Percent Increase in Mean Squared Error") + xlab("Variable Name") + theme_minimal() + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) + geom_hline(yintercept = 3.5, color ="blue") + coord_flip()
```

### Descriptive Statistics
```{r}
meansd = function(x) {
  average = round(mean(x, na.rm = T), 2)
  standard_dev = round(sd(x, na.rm = T), 2)
  paste0(average, "(", standard_dev, ")")
}
meansd(ENXadoption$adoption)
meansd(ENXadoption$bachelors_more)
meansd(ENXadoption$white)
meansd(ENXadoption$black)
meansd(ENXadoption$asian)
meansd(ENXadoption$hispanic)
meansd(ENXadoption$income100k_more)
meansd(ENXadoption$householdsize3_more)
meansd(ENXadoption$democrat)
meansd(ENXadoption$republican)
meansd(ENXadoption$unaffiliated_party)
meansd(ENXadoption$other_party)
meansd(ENXadoption$olderthan60)
meansd(ENXadoption$death_rate)
meansd(ENXadoption$maskfrequently_more)
meansd(ENXadoption$fully_vaccinated)
meansd(ENXadoption$residential)
meansd(ENXadoption$cases)
meansd(ENXadoption$hospitalizations)
meansd(ENXadoption$deaths)
```